---
title: "Section 1: Data Visualization & R"
author: '@laneharrison'
date: "2/20/2017"
output:
  ioslides_presentation: default
  beamer_presentation: default
  slidy_presentation: default
  html_document:
    keep_md: true
editor_options:
  chunk_output_type: console
---

## Data Visualization

Data Visualization is a key step in exploratory data analysis.

Visualization is also used to communicate results to others.

What we need is an *effective* and *flexible* means for producing charts in R.

## ggplot2 & The Grammar of Graphics

The [Grammar of Graphics]() is a systematic way of thinking about creating charts.

For our purposes, let's think about building a chart in _layers_.

## ggplot2 & The Grammar of Graphics

```{r}
library(tidyverse) # this loads library(ggplot2), with other useful libraries
ggplot()
```

## First, some data...

```{r}
mpg # note the variables
```

## Making a scatterplot

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))
```

## Let's check the documentation

- http://ggplot2.tidyverse.org/reference/geom_point.html
- The "cheatsheet" (built in!!!)

## Your turn!

Beginning with the code below, try:   

1. Displaying other variables 
2. Adding color 
3. (Advanced) Making a different chart (i.e. bar) 
```{r eval=FALSE}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))
```

## Bar

```{r}
ggplot(data = mpg) + 
  geom_bar(mapping = aes(x = class))
```



## Facets
One of the **most powerful** techniques for looking at complex data.

```{r eval=FALSE}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```

## Facets

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```

## Facet one variable against another

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_grid(drv ~ cyl)
```


## Going beyond

- Trend lines
- Density plots
- Interactivity (Shiny)
- Much much more

<!--
## Extra Credit
Re-do the examples with the dataset below
```{r eval=FALSE}
library('rio') #install.packages('rio')
vehicles <- rio::import("https://www.fueleconomy.gov/feg/epadata/vehicles.csv.zip")
vehicles
```
-->

